Exploring the Link Between Family Background and Children's Academic Achievement in the United States (2021)

### 2023-04-11 outputs/paper/paper.qmd
add the model for prestige score and add text to data cleaning

In my previous paper, there are several findings. The average years of schooling in the US is 14, which is higher than the global average of 8.7 years. A bachelor's degree and a graduate degree are more common than an associate or junior college degree, and the US ranks behind several countries in terms of higher education attainment for 25-34 year-olds. Parental education levels have a positive correlation, and there is a strong link between parents' education levels and their children's academic achievement and societal status. Fathers' education levels appear to have a stronger impact than mothers' on children's academic achievement, but other factors may also play a role. Further research is needed to fully understand the relationship between these variables. Policymakers can use this information to make informed decisions about resource allocation and policy
development. Here is a [link](https://github.com/ruibosun/how-parents-affect-childrens-education/blob/main/outputs/paper/paper.pdf) to view the entire paper.However, the prior study did not appropriately handle the issue of missing values (NAs). Consequently, this new paper aims to conduct a more rigorous analysis by appropriately addressing the NAs. Moreover, it will address the limitations of the previous study to produce more accurate and representative results.

@@

The data will be presented clearly and succinctly with plots and tables. This following analysis is processed in R [@R]  with packages of tidyverse [@tidyverse], dplyr [@dplyr], here [@here], haven [@haven] and broom [@broom]. The tables are constructed via knitr [@knitr], scales [@scales] and kableExtra [@kableExtra]. The package inside tidyverse helps to create the plots in ggplot2 [@tidyverse]. This paper is knitted as a PDF file by the packages of R markdown [@RMarkdown] and formatted using patchwork [@patchwork].

@@

The target population is US adults aged 18 and older. The GSS uses a multistage probability sampling approach to recruit its sample from the US Census Bureau's Master Address File. One strength of the GSS is its long history and large sample size, but a potential weakness is a reliance on self-reported data, which can be subject to response bias. The questionnaire includes both closed-ended and open-ended questions, but there may be concerns about biased or leading questions. The GSS employs various methods to adjust for non-response and ensure the representativeness of the sample. One strength of the GSS is its long history and large sample size, but a potential weakness is a reliance on self-reported data, which can be subject to response bias. The questionnaire includes both closed-ended and open-ended questions, but there may be concerns about biased or leading questions. [@norcuofch] The GSS employs various methods to adjust for non-response and ensure the representativeness of the sample. The limitation will be explained in detail in the @sec-weakness.

This paper will test and explore how parents’ social-economics status can affect children’s education. The data that will be used in this paper comes from the US General Social Survey from the National Opinion Research Center at the University of Chicago. The factor of socio-economics status is measured using the occupation prestige score. It is a measure used in social science research to assess the level of social status or prestige associated with a particular occupation or profession, which was developed by the National Opinion Research Center at the University of Chicago. As part of the GSS, respondents are asked to rate the prestige or social standing of various occupations on a scale of 1 to 90, with higher scores indicating greater prestige. [@smith] Another variable that has been used is the education level, which is a number of variables to indicate the number of years the respondents have spent in school and college. Other variables will be discussed in @sec-data.

@@

For simplicity purposes, only variables that are closely related to the topics are selected for further analysis. These variables are mapres10, papres10, paeduc, maeduc, educ, mawrkslf, pawrkslf, degree, born, sex, reg16, and incom16.

@@

None of the missing values (NA) is removed in this paper. The reason is that removing the missing values in every column would result in a huge loss in the dataset. To maintain the representativeness of the data, missing values are automatically filtered by R, rather than removed manually. The possible limitation and effect of the missing values will be discussed in the @sec-weakness.

>>>>

In my previous paper, there are several findings. The average years of schooling in the US is 14, which is higher than the global average of 8.7 years. A bachelor's degree and a graduate degree are more common than an associate or junior college degree, and the US ranks behind several countries in terms of higher education attainment for 25-34 year-olds. Parental education levels have a positive correlation, and there is a strong link between parents' education levels and their children's academic achievement and societal status. Fathers' education levels appear to have a stronger impact than mothers' on children's academic achievement, but other factors may also play a role. Further research is needed to fully understand the relationship between these variables. Policymakers can use this information to make informed decisions about resource allocation and policy
development. Here is a [link](https://github.com/ruibosun/how-parents-affect-childrens-education/blob/main/outputs/paper/paper.pdf) to view the entire paper. However, the prior study did not appropriately handle the issue of missing values (NAs). Consequently, this new paper aims to conduct a more rigorous analysis by appropriately addressing the NAs. Moreover, it will address the limitations of the previous study to produce more accurate and representative results.

@@

The data will be presented clearly and succinctly with plots and tables. This following analysis is processed in R [@R]  with packages of tidyverse [@tidyverse], dplyr [@dplyr], here [@here], haven [@haven] and broom [@broom]. The tables are constructed via knitr [@knitr], scales [@scales] and kableExtra [@kableExtra]. The package inside tidyverse helps to create the plots in ggplot2 [@tidyverse]. The missing values were imputed using the mice package [@mice]. This paper is knitted as a PDF file by the packages of R markdown [@RMarkdown] and formatted using patchwork [@patchwork].

@@

The target population is US adults aged 18 and older. The GSS uses a multistage probability sampling approach to recruit its sample from the US Census Bureau's Master Address File. One strength of the GSS is its long history and large sample size, but a potential weakness is a reliance on self-reported data, which can be subject to response bias. The questionnaire includes both closed-ended and open-ended questions, but there may be concerns about biased or leading questions. The GSS employs various methods to adjust for non-response and ensure the representativeness of the sample. The questionnaire includes both closed-ended and open-ended questions, but there may be concerns about biased or leading questions. [@norcuofch] The GSS employs various methods to adjust for non-response and ensure the representativeness of the sample. The limitation will be explained in detail in the @sec-weakness.

$$
Prestige_{ij}=\mu+\alpha_i+\beta_{j}+\epsilon_{ij}
$$ {#eq-modpr}

This paper will test and explore how parents’ social-economics status can affect children’s education. The data that will be used in this paper comes from the US General Social Survey from the National Opinion Research Center at the University of Chicago.

The factor of socio-economics status is measured using the occupation prestige score. It is a measure used in social science research to assess the level of social status or prestige associated with a particular occupation or profession, which was developed by the National Opinion Research Center at the University of Chicago. As part of the GSS, respondents are asked to rate the prestige or social standing of various occupations on a scale of 1 to 100, with higher scores indicating greater prestige. The ideal dataset would have 90,090 ratings from 1,001 raters, but due to various issues, only 82,800 ratings from 946 raters were used. Invalid or inconsistent ratings, such as cases with reversed codes, too few ratings, equal scores for all occupations, or a low standard deviation of ratings, were excluded to ensure the reliability of the occupational scores.[@smith]

Next,the prestige score is obtained by the @eq-modpr. $Prestige_{ij}$ is the prestige score for occupational title i and rater j. The $\alpha_i$ represents the occupational differences of interest, while the $\beta_j$ represents differences among the raters that need to be controlled for in estimating the $\alpha_i$. The $\mu$ is the overall mean prestige score, and $\epsilon_{ij}$ represents the random error term. The range of the prestige scores is from 0 to 100, with higher scores indicating a better or more prestigious occupation[@smith].

Another variable that has been used is the education level, which is a number of variables to indicate the number of years the respondents have spent in school and college. Other variables will be discussed in @sec-data.

@@

For simplicity purposes, only variables that are closely related to the topics are selected for further analysis. These variables are **mapres10**, **papres10**, **paeduc**, **maeduc**, **educ**, **degree**, **born**, **sex**, **reg16**, and **incom16**.

@@

As mentioned before, this paper is an revised version of [the previous paper](https://github.com/ruibosun/how-parents-affect-childrens-education/blob/main/outputs/paper/paper.pdf) which has some limitations and weaknesses. In order to improve the representiveness of the data, this paper use the multiple-imputation and dropping methods [@gelman]. There are two types of missing values: numercial and categorical missing values.

For numerical variables (**paeduc**, **maeduc**, **educ**, **mapres10** and **papres10**), multiple imputation techniques are employed to estimate and fill in the missing data points. This method generates a more accurate dataset without the missing values. Following the imputation, the resulting dataset is transformed into a new one with no NAs for further analysis.

Next, any categorical variables with missing values are addressed, such as **degree**, **sex**, **born**, **reg16** and **incom16**. In this case, all these variable with NAs are removed from the data, ensuring that the analysis is only performed on complete cases. This helps to minimize any biases that might arise due to incomplete data.
